\bf Forecasts Reconciliation: A geometric view with new insights on bias correction.

### 2019-08-06 Hierarchical-forecast-reconciliation/HierarchicalForecastReconciliation.tex

	The second property introduced above has been the focus of theoretical results in the forecast reconciliation literature.  He we provide a more streamlined version of proofs by \cite{VanErven2015a} and \cite{Wickramasuriya2017} before providing geometrical intuition aimed at simplifying the reader's understanding of these proofs.

	Consider the Euclidean distance between a forecast and the target. This is equivalent to the root of the sum of squared errors over the entire hierarchy. Let $\bm{y}_{t+h}$ be the realisation of the data generating process at time $t+h$, and let $\|\bm{v}\|_2$ be the $L_2$ norm of vector $\bm{v}$. The following theorem shows that reconciliation never increases, and in most cases reduces, the sum of squared errors of point forecasts.

	\begin{theo}[Distance reducing property]\label{th:distred}
		If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal projection onto $\mathfrak{s}$, then the following inequality holds:
		\begin{equation}
		\|(\tilde{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_2\le\|(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_2.
		\end{equation}
	\end{theo}
	\begin{proof}
		Since the aggregation constraints must hold for all realisations, $\bm{y}_{t+h}\in\mathfrak{s}$ and $\bm{y}_{t+h}=\bm{S}\bm{G}\bm{y}_{t+h}$ since $\bm{S}\bm{G}$ is a projection onto $\mathfrak{s}$. Therefore,
		\begin{align}
		\|(\tilde{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_2&=\|(\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}-\bm{S}\bm{G}\bm{y}_{t+h})\|_2\\
		&=\|\bm{S}\bm{G}(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_2.
		\end{align}
		The Cauchy-Schwarz inequality can be used to show that orthogonal projections are bounded operators \citep{Hun2001}, therefore
		\begin{equation*}
		\|\bm{S}\bm{G}(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_2\le
		\|(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_2.
		\end{equation*}
	\end{proof}
	The inequality is strict whenever $\hat{\bm{y}}_{t+h|t}\notin\mathfrak{s}$.

	\todo{compare to van Erven and Wickremasuriya more explictly}

	The simple geometric intuition behind the proof is demonstrated in Figure~\ref{fig:Schematic_OLSRecon}.  In this schematic, the coherent subspace is depicted as a black arrow.  The base forecast $\hat{\bm{y}}$ is shown as a blue dot.  Since $\hat{\bm{y}}$ is incoherent it does not lie in $\mathfrak{s}$.  Reconciliation is an orthogonal projection from $\hat{\bm{y}}$ to the coherent subspace yielding the reconciled forecast $\tilde{\bm{y}}$ shown in red.  Finally, the target of the forecast $\bm{y}$ is displayed as a black point, and although its exact location is unknown to the forecaster, it is known that it will lie somewhere along the coherent subspace.

@@

	Figure~\ref{fig:Schematic_OLSRecon} clearly shows that $\hat{\bm{y}}$, $\tilde{\bm{y}}$ and $\bm{y}$ form a right angled triangle with $\tilde{y}$ at the right-angled vertex.  In this triangle the line between $\bm{y}$ and $\hat{\bm{y}}$ is the hypotenuse and therefore must be longer than the distance between $\bm{y}$ and $\tilde{\bm{y}}$.  As such reconciliation is guaranteed to reduce the squared error of the forecast. \todo{Do we just include this as the proof rather than the existing proof to 3.1}

    Theorem~\ref{th:distred} is in some ways more powerful than perhaps previously understood.  It is often stated in terms of expectations\todo{check other proofs}.  However, the distance reducing property will hold for any realisation and any forecast and not just on average.  Nothing needs to be assumed about the statistical properties of the data generating process or the process by which forecasts are made.

@@

	\subsection{Oblique Projections}

@@

	Another way of understanding the generalised Euclidean geometry is that it is defined by the norm $\bm{v}'\bm{W}^{-1}\bm{v}$.  This interpretation is quite instructive when it comes to thinking about the connection between distances and loss functions.  In the generalised Euclidean geometry, the distance between the reconciled forecast and the realisation is given by $(\hat{\bm{y}}-\bm{y})'\bm{W}^{-1}(\hat{\bm{y}}-\bm{y})$.  For diagonal $\bm{W}$ this is equivalent to a weighted sum of squared error loss function and when $\bm{W}$ is a covariance matrix this is equivalent to a Mahalanobis distance.  As such Theorem~\ref{th:distred} can easily be generalised.  If the objective function is some weighted sum of squared errors, or a Mahalanobis distance, then the projection matrix $\bm{S}\left(\bm{S}'\bm{W}^{-1}\bm{S}\right)^{-1}\bm{S}'\bm{W}^{-1}$ is guaranteed to improve forecast accuracy over base forecasts, for an appropriately selected $\bm{W}$.

@@

	Figure~\ref{fig:MinT_justification1} provides geometrical intuition into the MinT method.  Suppose the in-sample errors are given by the orange points.  They provide information on the most likely direction of large deviations from the coherent subspace.  This direction is denoted by $\bm{R}$.  Figure~\ref{fig:MinT_justification2} then shows a target value of $\bm{y}$, while the grey points indicate possible values for the base forecasts (the base forecasts are of course stochastic).  One possible value of the forecast is depicted in blue as $\hat{\bm{y}}$.  An oblique projection of the blue point back along the direction of $\bm{R}$ yields a reconciled forecast closer to the target, especially compared to an orthogonal projection.  Imagining a similar oblique projection along $\bm{R}$ for all the grey points yield reconciled forecasts tightly packed near the target $\bm{y}$.  In this sense, the oblique MinT projection minimises the forecast error variance of reconciled forecasts.\todo{needs picture with all points projected obliquely then orthogonally}.

@@

	The property that projections map all vectors in the coherent subspace onto themselves is also useful in proving the unbiasedness preserving property of reconciliation \todo{reference Shanika and maybe van Erven Culigari}.  Before restating this proof using a  clear geometric interpretation we discuss in a precise fashion what is meant by unbiasedness.

>>>>

	The property that reconciliation should improve forecasts was touched upon in Section 2.3 of \cite{Wickramasuriya2017}.  The discussion in that paper focuses on the case of MinT. Here we provide a new explicit proof of that result first in the case of an orthogonal projection where the geometric intuition of the proof is clear and then generalise the result to reconciliation by any projection matrix in Section~\ref{sec:oblique}.

	Consider the Euclidean distance between a forecast and the target. This is equivalent to the root of the sum of squared errors over the entire hierarchy. Let $\bm{y}_{t+h}$ be the realisation of the data generating process at time $t+h$. The following theorem shows that reconciliation never increases, and in most cases reduces, the sum of squared errors of point forecasts.

	\begin{theo}[Distance reducing property]\label{th:distred}
		If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal (in the Euclidean sense) projection onto $\mathfrak{s}$ and let $\|\bm{v}\|$ be the $L_2$ norm (in the Euclidean sense) of vector $\bm{v}$ then:
		\begin{equation}
		\|(\tilde{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|\le\|(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|.
		\end{equation}
	\end{theo}
	\begin{proof}
		Since, $\bm{y}_{t+h},\tilde{\bm{y}}_{t+h}\in\mathfrak{s}$ and since the projection is orthogonal, by Pythagoras' theorem
		\begin{equation}
		\|(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|^2=\|(\tilde{\bm{y}}_{t+h|t}-\hat{\bm{y}}_{t+h})\|^2+\|(\tilde{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|^2.
		\end{equation}

		Since $\|(\tilde{\bm{y}}_{t+h|t}-\hat{\bm{y}}_{t+h})\|^2\ge 0$ this implies,
		\begin{equation}
		\|(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|^2\ge\|(\tilde{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|^2.
		\end{equation}
		with equality only holding when $\tilde{\bm{y}}_{t+h|t}=\hat{\bm{y}}_{t+h}$.  Taking the square root of both sides proves the desired result.
	\end{proof}

	The simple geometric intuition behind the proof is demonstrated in Figure~\ref{fig:Schematic_OLSRecon}.  In this schematic, the coherent subspace is depicted as a black arrow.  The base forecast $\hat{\bm{y}}$ is shown as a blue dot.  Since $\hat{\bm{y}}$ is incoherent, $\hat{\bm{y}}_{t+h|t}\notin\mathfrak{s}$ and in this case the inequality is strict.  Reconciliation is an orthogonal projection from $\hat{\bm{y}}$ to the coherent subspace yielding the reconciled forecast $\tilde{\bm{y}}$ shown in red.  Finally, the target of the forecast $\bm{y}$ is displayed as a black point, and although its exact location is unknown to the forecaster, it is known that it will lie somewhere along the coherent subspace.

@@

	Figure~\ref{fig:Schematic_OLSRecon} clearly shows that $\hat{\bm{y}}$, $\tilde{\bm{y}}$ and $\bm{y}$ form a right angled triangle with $\tilde{y}$ at the right-angled vertex.  In this triangle the line between $\bm{y}$ and $\hat{\bm{y}}$ is the hypotenuse and therefore must be longer than the distance between $\bm{y}$ and $\tilde{\bm{y}}$.  As such reconciliation is guaranteed to reduce the squared error of the forecast.

    Theorem~\ref{th:distred} is in some ways more powerful than perhaps previously understood.  Crucially, the result is not a result that requires taking expectations.  This distance reducing property will hold for any realisation and any forecast and not just on average.  Nothing needs to be assumed about the statistical properties of the data generating process or the process by which forecasts are made. \todo{include evidence from small empirical example.}

@@

	\subsection{Oblique Projections}\label{sec:oblique}

@@

	Another way of understanding the generalised Euclidean geometry is that it is defined by the norm $\bm{v}'\bm{W}^{-1}\bm{v}$.  This interpretation is quite instructive when it comes to thinking about the connection between distances and loss functions.  In the generalised Euclidean geometry, the distance between the reconciled forecast and the realisation is given by $(\hat{\bm{y}}-\bm{y})'\bm{W}^{-1}(\hat{\bm{y}}-\bm{y})$.  For diagonal $\bm{W}$, this is equivalent to a weighted sum of squared error loss function and when $\bm{W}$ is a covariance matrix, this is equivalent to a Mahalanobis distance.  As such Theorem~\ref{th:distred} can easily be generalised as follows:

	\begin{theo}[General distance reducing property]\label{th:gdistred}
        If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal (in the generalised Euclidean sense) projection onto $\mathfrak{s}$ and let $\|\bm{v}\|_{\bm{W}}$ be the $L_2$ norm (in the generalised Euclidean sense with respect to $\bm{W}$) of vector $\bm{v}$ then:
		\begin{equation}
		\|(\tilde{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_{\bm W}\le\|(\hat{\bm{y}}_{t+h|t}-\bm{y}_{t+h})\|_{\bm W}.
		\end{equation}
	\end{theo}
	\begin{proof}
    The proof is identical to the proof for Theorem~\ref{th:distred} but relies on the generalised Pythagorean Theorem (applicable to Generalised Euclidean space) rather than the Pythagorean Theorem.
    \end{proof}

	The implication of Theorem~\ref{th:gdistred} is that if the objective function is some weighted sum of squared errors, or a Mahalanobis distance, then the projection matrix $\bm{S}\left(\bm{S}'\bm{W}^{-1}\bm{S}\right)^{-1}\bm{S}'\bm{W}^{-1}$ is guaranteed to improve forecast accuracy over base forecasts, for an appropriately selected $\bm{W}$.

	Note here that we rely here on the generalised Pythagorean Theorem (which involves an equality).  In contrast, \cite{Wickramasuriya2017} follow \cite{VanErven2015a} in stating their result in terms of the Generalised Pythagorean Inequality.  The proof of \cite{Wickramasuriya2017} requires an assumptions about convexity so that the angle between the base forecast and coherent subspace must be greater than 90 degrees.  The proof we have provided here requires no such assumption, since this may not hold for an arbitrary $\bm{W}$.  As such the statement from \cite{Wickramasuriya2017} that {\em``MinT reconciled forecasts are at least as good as the incoherent forecasts''} should be qualified; this is only true in with respect to a loss function that depends on ${\bm W}$.  If Euclidean distance (or mean squared error) is used, there will be cases where the MinT estimator does not improve upon base forecasts\todo{include empirical example}.

@@

	Figure~\ref{fig:MinT_justification1} provides geometrical intuition into the MinT method.  Suppose the in-sample errors are given by the orange points.  They provide information on the most likely direction of large deviations from the coherent subspace.  This direction is denoted by $\bm{R}$.  Figure~\ref{fig:MinT_justification2} then shows a target value of $\bm{y}$, while the grey points indicate possible values for the base forecasts (the base forecasts are of course stochastic).  One possible value of the forecast is depicted in blue as $\hat{\bm{y}}$.  An oblique projection of the blue point back along the direction of $\bm{R}$ yields a reconciled forecast closer to the target, especially compared to an orthogonal projection.  Imagining a similar oblique projection along $\bm{R}$ for all the grey points yield reconciled forecasts tightly packed near the target $\bm{y}$.  In this sense, the oblique MinT projection minimises the forecast error variance of reconciled forecasts.\todo{needs picture with all points projected obliquely then orthogonally}.    In contrast to the result in Theorem~\ref{th:gdistred}, this property is a statistical property in the sense that MinT is optimal is expectation.

@@

	The property that projections map all vectors in the coherent subspace onto themselves is also useful in proving the unbiasedness preserving property of reconciliation \cite{Wickramasuriya2017}.  Before restating this proof using a  clear geometric interpretation we discuss in a precise fashion what is meant by unbiasedness.
